### 2020-02-21 tex/preface.tex

\begin{pre}
	\thispagestyle{empty}
\begin{center}
  {\kaishu{材料来自于网上，如有侵权，请联系ocnzhao@163.com予以纠正,THX}}
\end{center}
	\begin{center}
%		{\kaishu{人在春风和气中}}
\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.75\textwidth]{Imagetreemath.png}
\end{figure}
	\end{center}

%\vspace*{5\baselineskip}
%\centerline{\includegraphics[scale=0.6]{example/gzh.jpg}}
%\centerline{\fontsize{26pt}{26pt} 微信公众号}
\end{pre}